Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-13 paper.tex
Another example of an MDP

@@

% TODO: short intro on MDPs with references to the literature, why they are
% important
% TODO: turn the door into an actual door

MDPs have been used to solve a wide range of problems: optimising airline yield
management \cite{DBLP:journals/transci/LautenbacherS99}, constructing optimal
dialogue strategies \cite{658989,DBLP:conf/icassp/LevinPE98}, electricity supply
bidding \cite{867150}, and many others.

@@

% TODO: minimal example with a diagram

@@

\begin{example} \label{example}
  Consider an MDP $(S, \overline{s}, A, P, L)$, where $S = \{ s_0, s_1, s_2, s_3
  \}$, $\overline{s} = s_0$, $A = \{ a, b, c \}$, and $P, L$ are defined as
  follows:
  \begin{equation*}
    \begin{split}
      P(s_0, a) &= [s_1 \mapsto 1], \\
      P(s_1, b) &= [s_0 \mapsto 0.7, s_1 \mapsto 0.3], \\
      P(s_1, c) &= [s_2 \mapsto 0.5, s_3 \mapsto 0.5], \\
      P(s_2, a) &= [s_2 \mapsto 1], \\
      P(s_3, a) &= [s_3 \mapsto 1],
    \end{split}
    \qquad
    \begin{split}
      L(s_0) &= \{ initial \}, \\
      L(s_1) &= \emptyset, \\
      L(s_2) &= \{ heads \}, \\
      L(s_3) &= \{ tails \}.
    \end{split}
  \end{equation*}
  Furthermore, equip it with a reward structure $(\rho, \iota)$, where
  $\rho(s_2) = 3$, $\iota(s_1, b) = 1$, and both functions are zero everywhere
  else.
\end{example}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    % outside walls
    \draw (3, 2) -- (3, 0) -- (1, 0) -- (1, 2) -- (0, 2) -- (0, 3) -- (3, 3) -- (3, 2) -- (2, 2);
    % the probabilistic wall
    \draw[red,ultra thick,dashed] (2, 2) -- (2, 1);
    % the background grid
    \draw[gray,very thin] (1, 3) -- (1, 2);
    \draw[gray,very thin] (2, 3) -- (2, 2);
    \draw[gray,very thin] (1, 1) -- (3, 1);
    \draw[gray,very thin] (2, 1) -- (2, 0);
    % the agent
    \node at (0.5, 2.3) {\PHchild};
    % the door
    \draw[dotted] (1.25,2) arc (180:270:0.5cm);
    \draw[brown,ultra thick] (1.75,1.5) -- (1.75,2) node[right,  near start]{};
    \draw (1, 2) -- (1.25, 2);
    \draw (1.75, 2) -- (2, 2);
  \end{tikzpicture}
  \caption{The full map. The agent is in the top left cell. There is a thick
    brown door connecting the middle cell of the top row with a $2 \times 2$
    room below. The red dashed wall exists with probability $1-p$.}
  \label{agent2_map}
\end{figure}

>>>>

\usetikzlibrary{shapes,arrows.meta}

@@

Markov decision processes \cite{bellman1957markovian,howard1960dynamic} provide
a way to model decision making in situations with both probabilistic outcomes
and nondeterministic decision making. They have been used to solve a wide range
of problems: optimising airline yield management
\cite{DBLP:journals/transci/LautenbacherS99}, constructing optimal dialogue
strategies \cite{658989,DBLP:conf/icassp/LevinPE98}, electricity supply bidding
\cite{867150}, and many others
\cite{DBLP:journals/questa/StidhamW93,white1985real,white1988further,white1993survey}.
Furthermore, MDPs have been extended in multiple ways, each extension bringing
its own applications: constrained MDPs \cite{altman1999constrained}, partially
observable MDPs \cite{astrom1965optimal}, fuzzy MDPs \cite{fakoor2016humanoid},
and continuous-time MDPs \cite{guo2009continuous}.

@@

\begin{example}
  Let $M = (S, \overline{s}, A, P, L)$ be an MDP, where $S = \{ s_0, s_1, s_2,
  s_3 \}$, $\overline{s} = s_0$, $A = \{ \text{start}, \text{wait}, \text{send},
  \text{restart}, \text{stop} \}$, and $P, L$ are defined as follows:
  \begin{equation*}
    \begin{split}
      P(s_0, \text{start}) &= P(s_1, \text{wait}) = [s_1 \mapsto 1], \\
      P(s_1, \text{send}) &= [s_2 \mapsto 0.01, s_3 \mapsto 0.99], \\
      P(s_2, \text{restart}) &= [s_0 \mapsto 1], \\
      P(s_3, \text{stop}) &= [s_3 \mapsto 1],
    \end{split}
    \qquad
    \begin{split}
      L(s_0) &= \emptyset, \\
      L(s_1) &= \{ \text{try} \}, \\
      L(s_2) &= \{ \text{failure} \}, \\
      L(s_3) &= \{ \text{success} \}.
    \end{split}
  \end{equation*}
  Then $M$ can represent a simple communication protocol. The process
  initialises, and is faced with a nondeterministic choice between waiting and
  sending a message. Most of the time the sent message is successfully
  delivered, and the process stops. But sometimes a message fails to deliver, in
  which case the process has to restart. See Fig. \ref{first_mdp} for a
  visualisation of $M$.
\end{example}

\begin{figure}
  \centering
  \begin{tikzpicture}[state/.style={rectangle, draw, scale=2},
    action/.style={diamond, draw, fill=lightgray, minimum size=1.6cm}]
    \node(origin) at (-1, 0){};
    \node[state](s0) at (0, 0){$s_0$};
    \node[action](start) at (2, 0){start};
    \node[state,label=try](s1) at (4, 0){$s_1$};
    \node[action](wait) at (4, -2){wait};
    \node[action](send) at (6, 0){send};
    \node[state,label={right:failure}](s2) at (8, 2){$s_2$};
    \node[action](restart) at (8, 4){restart};
    \node[state,label={below:success}](s3) at (8, -2){$s_3$};
    \node[action](stop) at (10, -2){stop};
    \draw[-{>[scale=2]}] (s0) edge (start);
    \draw[-{>[scale=2]}] (start) edge node[above] {$1$} (s1);
    \draw[-{>[scale=2]},bend left] (s1) edge (wait);
    \draw[-{>[scale=2]},bend left] (wait) edge node[left] {$1$} (s1);
    \draw[-{>[scale=2]}] (s1) edge (send);
    \draw[-{>[scale=2]}] (send) edge node[left] {$0.01$} (s2);
    \draw[-{>[scale=2]}] (send) edge node[left] {$0.99$} (s3);
    \draw[-{>[scale=2]}] (s2) edge (restart);
    \draw[-{>[scale=2]},bend right] (restart) edge node[above] {$1$} (s0);
    \draw[-{>[scale=2]},bend left] (s3) edge (stop);
    \draw[-{>[scale=2]},bend left] (stop) edge node[below] {$1$} (s3);
    \draw[-{>[scale=2]}] (origin) edge (s0);
  \end{tikzpicture}
  \caption{A simple MDP for communication.}
  \label{first_mdp}
\end{figure}

@@

\begin{example} \label{example}
  Consider an MDP $(S, \overline{s}, A, P, L)$, where $S = \{ s_0, s_1, s_2, s_3
  \}$, $\overline{s} = s_0$, $A = \{ a, b, c \}$, and $P, L$ are defined as
  follows:
  \begin{equation*}
    \begin{split}
      P(s_0, a) &= [s_1 \mapsto 1], \\
      P(s_1, b) &= [s_0 \mapsto 0.7, s_1 \mapsto 0.3], \\
      P(s_1, c) &= [s_2 \mapsto 0.5, s_3 \mapsto 0.5], \\
      P(s_2, a) &= [s_2 \mapsto 1], \\
      P(s_3, a) &= [s_3 \mapsto 1],
    \end{split}
    \qquad
    \begin{split}
      L(s_0) &= \{ \text{initial} \}, \\
      L(s_1) &= \emptyset, \\
      L(s_2) &= \{ \text{heads} \}, \\
      L(s_3) &= \{ \text{tails} \}.
    \end{split}
  \end{equation*}
  Furthermore, equip it with a reward structure $(\rho, \iota)$, where
  $\rho(s_2) = 3$, $\iota(s_1, b) = 1$, and both functions are zero everywhere
  else.
\end{example}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    % outside walls
    \draw (3, 2) -- (3, 0) -- (1, 0) -- (1, 2) -- (0, 2) -- (0, 3) -- (3, 3) -- (3, 2) -- (2, 2);
    % the probabilistic wall
    \draw[red,ultra thick,dashed] (2, 2) -- (2, 1);
    % the background grid
    \draw[gray,very thin] (1, 3) -- (1, 2);
    \draw[gray,very thin] (2, 3) -- (2, 2);
    \draw[gray,very thin] (1, 1) -- (3, 1);
    \draw[gray,very thin] (2, 1) -- (2, 0);
    % the agent
    \node at (0.5, 2.3) {\PHchild};
    % the door
    \draw[dotted] (1.25,2) arc (180:270:0.5cm);
    \draw[brown,ultra thick] (1.75,1.5) -- (1.75,2) node[right,  near start]{};
    \draw (1, 2) -- (1.25, 2);
    \draw (1.75, 2) -- (2, 2);
  \end{tikzpicture}
  \caption{The full map. The agent is in the top left cell. There is a door
    connecting the middle cell of the top row with a $2 \times 2$ room below.
    The dashed wall exists with probability $1-p$.}
  \label{agent2_map}
\end{figure}
